---
layout: post
title: "Basic Data Visualization in R"
author: "Minjeong Kim"
date: "2023-02-22"
output:
  md_document:
    variant: markdown_github
    preserve_yaml: true
---
<style type="text/css">

body, td {
   font-family: times, serif; font-size: 14px;
}
code.r{
  font-family: times, serif; font-size: 12px;
}
pre {
  font-family: times, serif; font-size: 12px
}
</style>

**Data Visualization in R**

This is a lecture note for a data science course ("Data Science for Psychology Major") at San Francisco State University (lectured by Dr. Gaurav Suri). This lecture was intended to learn different methods of data visualization to gather graphical interpretation on given data.

Data Used: mpg, Orange, swiss, a self-created data frame
Package Used: tidyverse

```{r}
knitr::opts_knit$set(base.dir = "D:/MJ/study/Personal Project/Github Portfolio/Github Blog/mj-kim904.github.io/_posts", base.url = "/")
knitr::opts_chunk$set(fig.path = "images/")
```


```{r load necessary package}
library(tidyverse)
```

```{r}
head(mpg)
# The head() function returns truncated rows.

summary(mpg)
# The summary() function tells you the various fields and summarizes the quantitative fields.
```

Here is the code for visualization using the ggplot() function.

[Code]
ggplot(data = <DATA>) + <GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))


**Scatter Plot**

[Code]
ggplot(data = <DATA>) + geom_point(mapping = aes(<MAPPINGS>))

[Example]
```{r}
#Highway Miles per Gallon vs. Engine Displacement (in litre)
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

#Highway Miles per Gallon vs. City Miles per Gallon
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = cty, y = hwy))

```


**Bar Graph**

```{r Sample data creation}
survey <- data.frame(fruit=c("Apple", "Banana", "Grapes", "Kiwi", "Orange", "Pears"), people=c(40, 50, 30, 15, 35, 20))
survey
```

[Code]
ggplot(data = <DATA>, aes(<MAPPINGS>)) + geom_bar(stat = "identity")

[Example]

```{r}
ggplot(data = survey, aes(x=fruit, y=people)) +
  geom_bar(stat = "identity")

#The argument "stat = identity" is telling it not to transform the data.
```

Additional Arguments

```{r}
#Make bars with color
ggplot(survey, aes(x=fruit, y=people, fill=fruit)) + 
  geom_bar(stat="identity")

#Make bars with specific color
ggplot(survey, aes(x=fruit, y=people, fill=fruit)) + 
  geom_bar(stat="identity") +
  scale_fill_manual(values=c("red2", "yellow2", "slateblue4", "green3", "orange", "olivedrab2"))

#Make bars with grey scale
ggplot(survey, aes(x=fruit, y=people, fill=fruit)) + 
  geom_bar(stat="identity") +
  scale_fill_grey()

#Make it in different theme
ggplot(survey, aes(x=fruit, y=people, fill=fruit)) + 
  scale_fill_manual(values=c("red2", "yellow2", "slateblue4", "green3", "orange", "olivedrab2")) +
  geom_bar(stat="identity") +
  theme_classic()

#Modify bar width
ggplot(survey, aes(x=fruit, y=people, fill=fruit)) + 
  scale_fill_manual(values=c("red2", "yellow2", "slateblue4", "green3", "orange", "olivedrab2")) +
  geom_bar(stat="identity", width = 0.5) +
  theme_classic()

#Add titiles and label axes
ggplot(survey, aes(x=fruit, y=people, fill=fruit)) + 
  scale_fill_manual(values=c("red2", "yellow2", "slateblue4", "green3", "orange", "olivedrab2")) +
  geom_bar(stat="identity", width = 0.5) +
  theme_classic() +
  ggtitle("Favorite fruit survey") +
  xlab("Fruits") +
  ylab("Number of People")

#add labels 

ggplot(survey, aes(x=fruit, y=people, fill=fruit)) + 
  scale_fill_manual(values=c("red2", "yellow2", "slateblue4", "green3", "orange", "olivedrab2")) +
  geom_bar(stat="identity", width = 0.5) +
  geom_text(aes(label=people), vjust=-0.3, size=3.5) +
  theme_classic() +
  ggtitle("Favorite fruit survey") +
  xlab("Fruits") +
  ylab("Number of People")

```

**Line Graph**

Here, we will use a different data set called "Orange".

[Code]
ggplot(data = <DATA>) + geom_line(mapping = aes(<MAPPINGS>))

[Example]
```{r}
#Circumference vs. Age
ggplot(data = Orange) +
  geom_line(mapping = aes(x = age, y = circumference, color = Tree))

```

Now, we will create a line graph for tree 1 only.

```{r}
head(Orange)

#Select the data for tree 1 and create a new data frame with it
tree_1 <- Orange[which(Orange$Tree==1),]
tree_1
```

```{r}
ggplot(data = tree_1) +
  geom_line(mapping = aes(x = age, y = circumference))
```


Additional Arguments

```{r}
#Give color to the line
ggplot(data = tree_1) +
  geom_line(mapping = aes(x = age, y = circumference), color = "blue") +
  theme_classic() +
  ggtitle("Tree 1: Circumference vs. Age") +
  xlab("Age") +
  ylab("Circumference")
```


**Scatter Plot with Line Added**

In the section, we will use "swiss" data set and learn how to add a line to a scatter plot.

This data set shows the standardized fertility measure and socio-economic indicators for each of 47 French-speaking provinces of Switzerland at about 1888. 

```{r Swiss data}
swiss
head(swiss)
summary(swiss)
str(swiss)

plot(swiss)
```


First, we will use basic function called "plot()" to graph.

[Code]
plot(<MAPPINGS>)
abline(lm(MAPPINGS))
- For the mapping of plot() function, x variable comes first followed by y variable.
- For the mapping of abline() function, y variable comes first followed by x variable.

[Example]

```{r plot function with line}
#1) Education vs. Fertility
plot(swiss$Fertility, swiss$Education)
#Add line
abline(lm(swiss$Education ~ swiss$Fertility))

#2) Fertility vs. Education
plot(swiss$Education, swiss$Fertility)
#Add line
abline(lm(swiss$Fertility ~ swiss$Education))
```


Now we will create the same plot using ggplot()

[Code]
coef(lm(<VAR2>~<VAR1>, data = <DATA>))

ggplot(data = <DATA>) + 
geom_point(mapping = aes(<MAPPINGS>)) +
geom_abline(slope = <SLOPE>, intercept = <INTERCEPT>)

[Example]

```{r}
ggplot(data = swiss) + 
  geom_point(mapping = aes(x = Education, y = Fertility))

#Find slope and intercept using coef() function
coef(lm(Fertility ~ Education, data = swiss))

ggplot(data = swiss) + 
  geom_point(mapping = aes(x = Education, y = Fertility)) +
  geom_abline(slope = -0.862, intercept = 79.61)
```

*More information about the line can be found in the posting for the linear regression. 

